Requantization of time-dependent mean field for pairing collective motion in superfluid nuclei

### 2018-11-20 thesis/thesis.tex
thesis

\usepackage{graphicx} % support the \includegraphics command and options

@@

\tableofcontents

\chapter{Introduction}

\section{Pairing interaction in nuclei}

Pairing correlation plays an important role near the Fermi energy. Pairing correlation is the two-body interaction which couples two identical nucleons into $J^{\pi}=0^+$ state. There are many evidences of pairing correlation detected from experiment. The most clear evidence is shown in Fig. \ref{Sn_isotope}. All of the ground states of even-even Sn isotopes are $J^{\pi}=0^+$ states  It indicates that the ground states consist of the $J^{\pi}=0^+$ pairs is more stable than other configurations. In addition, the ground states between even-even nuclei and neighborhood odd nuclei has large gaps of binding energy. In odd nuclei, the unpaired last neutron is the last single-particle level. The odd-even mass difference  To break the $J^{\pi}=0^+$ pair, we need large energy which is about $2\Delta \approx 24A^{-1/2}$ MeV.
\begin{figure}[htbp]
 \begin{center}
    \includegraphics[width=100mm, bb=0 0 400 300]{images/Sn_isotope.png}
 \end{center}
  \caption{Low-lying excited states in Sn isotopes \cite{}. The absolute values of binding energy are adjusted at 0 MeV in the ground state of ${}^{116}$Sn.}
  \label{Sn_isotope}
\end{figure}

\subsection{A subsection}

More text.

@@

To examine the pairing dynamics influencing the structure of nuclei, we can only concentrate the nucleons near the Fermi energy. The Hamiltonian of pairing model is
\begin{align}
	H &= \sum_{l=1}^L \epsilon_l n_l - g \sum_{l,l'} S_l^+ S_{l'}^- ,
\end{align}
where
\begin{align}
	n_l &= \sum_m a^{\dag}_{lm}a_{lm} \\
        S_l^{+} &= \sum_{m>0}a_{lm}^{\dag}a_{l\overline{m}}^{\dag} ,
\quad   S_l^{-} = S_l^{+\dag} .
\label{S_+}
\end{align}
In $L$ single-particle level system, each single-particle energy $\epsilon_l$ possesses $(2\Omega_l)$-fold
degeneracy ($\Omega_l=j_l+1/2$)
and $\sum_{m>0}$ indicates the summation over $m=1/2,3/2,\cdots,$ and $\Omega_l-1/2$. $\overline{m}$ indicates the time reversal quantum number with respect to $m$. The strength of pairing correlation is represented by coupling constant $g$. We also define a seniority $\nu_l$, which is the number of unpaired particles in each level. The unpaired particles have no pairing interaction, and play a part in Pauli blocking effect.\par
In one-level system ($L=1$), the pairing dynamics is well understood. The analytic solution of the total energy is
\begin{align}
	E_{\nu}(N) = \epsilon N - \frac{1}{4}g(N-\nu)(2\Omega-N-\nu+2) .
	\label{onelevel}
\end{align}
The pairing rotational band created by the pairing correlation is completely parabolic with the moment of inertia $2/g$. Also, the excited states in a specified $N$ is only described by the quantum number $\nu$. For example, the first excited state in even particle system is $\nu=2$ state, which correspond to one $J^{\pi}=0^+$ pair broken state.\par
The profound dynamics occurs in multi-level pairing model ($L\ge 2$). The pairing rotational mode becomes complex due to the shell effect. In addition, several pairing vibrational modes emerge. The pairing vibration is not single particle motion specified by $\nu_l$, but a collective motion with the excitation of $J^{\pi}=0^+$ pairs. In this chapter, we explain the numerical exact solution and derive the TDHFB dynamics in multi-level pairing Hamiltonian.

@@

\subsection{Richardson equation}
As mentioned above, diagonalizing the Hamiltonian is simple method to obtain the exact solution.
However, diagonalization is time consuming when the model space becomes large. The calculation time is proportional to the cubes of the dimension of the model space For example, there are 16 single particle levels between the magic number 50 and 82 in deformed nuclei. Because the dimension for the system is $D=2^{16}=65536$, the calculation time in deformed system is $\sim 10^5$ times as much as the case in spherical system. \par
The alternative method to obtain the exact solution is to solve Richardson equation. The idea was introduced by R. W. Richardson \cite{Richardson1, Richardson2, Richardson3}. We divide the total energy into two parts
\begin{align}
  E &= \braket{\psi|H|\psi} \\
  &= \sum_l \epsilon_l \nu_l + \sum_{\alpha=1}^M E_{\alpha} .
\end{align}
The first term is the energy of unpaird particle, and the second term is the total energy of $J^{\pi}=0^+$ pairs. The pair-energy $E_{\alpha}$ is derived from the non-linear equations
\begin{equation}
  1 - g\sum_l \frac{\Omega_l-\nu_l}{2\epsilon_l-E_{\alpha}} + 2g\sum_{\beta(\neq\alpha)}^M \frac{1}{E_{\beta}-E_{\alpha}} = 0 .
\end{equation}
$M$ indicates the number of $J^{\pi}=0^+$ pairs in the system, and equals to the numbers of the non-linear equations. $E_{\alpha}$ can become not only real numbers, but also complex conjugate pairs when $M\ge 2$.

\section{TDHFB dynamics}
%In realistic nuclear system with nuclear effective interaction, the necessary model space is huge. To diagonalize the effective Hamiltonian is extremely difficult task. Such study is done by shell model calculation. In alternative way, we only prepare a limited model space, which supposed to be reproduce the important parts of nuclear properties. The time-dependent mean-field theory 
To understand the classical picture of the pairing dynamics, time-dependent mean-field (TDMF) theory is useful tool. With pairing correlation, TDMF theory attributes to TDHFB theory.
The TDHFB equation can be derived from the time-dependent variational principle,
\begin{equation}
  \delta\mathcal{S}=0, \quad 
  \mathcal{S} \equiv \int \braket{\phi(t)|i\frac{\partial}{\partial t}-H|\phi(t)}dt = 0 ,
\end{equation}
where $\ket{\phi(t)}$ is the time-dependent generalized Slater determinant (coherent state) obtained from Thouless' theorem
\begin{equation}
  \ket{\phi(t)} = \mathcal{R}\exp{\left( \sum_{k<k'}Z_{kk'}(t)\beta_k^{\dag}\beta_{k'}^{\dag}
    \right)}\ket{\Phi_0} .
\end{equation}
$\beta_k^{\dag}$ is quasiparticle creation operator and $\mathcal{R}$ is normalization factor. $\ket{\Phi_0}$ is generally chosen as ground state or vacuum state. $Z_{kk'}(t)$ is time-dependent complex number determined by TDHFB equation.

@@

\subsection{Constant of motion}
\label{constant}
The number of the TDHFB degrees of freedom is same as the number of the single-particle levels. However, because the Hamiltonian depends only on the relative difference in the gauge angles, the global gauge angle 
\begin{align}
	\Phi = \frac{1}{L}\sum_l \chi_l
\end{align}
is a cyclic variable. With the definition for relative gauge angle (e.g. $\phi_l = \chi_l-\chi_1$), we can find that the conjugate momentum of $\Phi$ is the sum of $j_l$
\begin{align}
	J \equiv \frac{\partial\mathcal{L}}{\partial\dot{\Phi}} = \sum_l j_l .
\end{align}
From (\ref{occ}), $J$ is related to the total particle number
\begin{align}
	N = \sum_l \nu_l + 2J .
\end{align}
The Hamilton's equation for $J$
\begin{align}
	\dot{J} = -\frac{\partial\mathcal{H}}{\partial \dot{\Phi}} = 0, \quad J = {\rm const}
\end{align}
indicates that the total particle number is conserved in time evolution.
The canonical variable $(\Phi,J)$ describes the pairing rotation. The pairing rotational band emerges in the particle number chain ($\cdots,N-2,N,N+2,\cdots$). The last term in (\ref{Hamiltonian}) contains the higher order terms with respect to $N$. It related to the shell effect. In single-level system, the last term in (\ref{Hamiltonian}) disappears and (\ref{Hamiltonian}) attributes to (\ref{onelevel}). The pairing rotational band is exactly parabolic. In multi-level system, the coupling constant $g$ and single-particle energy $\epsilon_l$ affects significantly to the last term in (\ref{Hamiltonian}). The parabolic structure can be broken when the shell effect is strong.

@@

\section{Derivation of the collective subspace in adiabatic self-consistent collective coordinate method}

\section{Application of SPA in non-integrable system}

@@

\section{Properties of the spin-coherent state}
\label{formula}
The time-dependent coherent state classified into the spin-coherent state is employed in (\ref{coherent}). We derive the necessary formula for the spin-coherent state in this subsection. \par
First, we consider the single-level case.
With the commutation relation in (\ref{SU2commu}), the SU(2) quasispin operators fulfill the following relations
\begin{align}
  \hat{S}^{\pm}\ket{S,S^0} &= \sqrt{(S\mp S^0)(S\pm S^0+1)} \ket{S,S^0\pm 1}
  \label{S_pm} \\
  \hat{S}^0\ket{S,S^0} &= S^0\ket{S,S^0} .
  \label{S_0}
\end{align}
Using above relations, the coherent state can be expanded as 
\begin{align}
  \ket{Z} &= (1+|Z|^2)^{-S}e^{Z \hat{S}^+}\ket{S,S^0=-S} \nonumber \\
  &= (1+|Z|^2)^{-S}\sum_{n=0}^{2S}\sqrt{\frac{(2S)!}{n!(2S-n)!}}Z^n\ket{S,-S+n} .
\end{align}
We can find the normalization condition is fulfilled
\begin{align}
  \braket{Z|Z} &= (1+|Z|^2)^{-2S}\sum_{n=0}^{2S}\frac{(2S)!}{n!(2S-n)!}|Z|^{2n} \nonumber \\
  &= (1+|Z|^2)^{-2S}\sum_{n=0}^{2S} {}_{2S}\mathrm{C}_n(|Z|^2)^n \nonumber \\
  &= (1+|Z|^2)^{-2S}\times(1+|Z|^2)^{2S} \nonumber \\
  &= 1 .
\end{align}
The overlap is
\begin{align}
  \braket{\eta|Z} &= \frac{(1+\eta^*Z)^{2S}}{(1+|\eta|^2)^S(1+|Z|^2)^S} .
\end{align}

>>>>

\usepackage[dvips]{graphicx} % support the \includegraphics command and options
\usepackage[dvips]{color}

@@

\pagenumbering{roman}
\tableofcontents

\chapter{Introduction}
\pagenumbering{arabic}

\section{Pairing correlation in theoretical and experimental aspects}

\subsection{Superfluid nuclei}

Mean field model is standard tool to describe the structure of atomic nuclei. The idea of mean field model is based on the primary theoretical framework of shell model introduced by Mayer and Jensen in 1949. Mean field picture indicates that the nucleon can be described in single-particle motion in the mean field potential created by the whole nucleons in the nucleus. Although the many-body correlation between the nucleons is extremely complicate, the single-particle motion picture can well reproduce the basic properties (e.g. radius, binding energy, separation energy, magic number) in most of nuclide. %%The most popular mean field potential is Woods-Saxon potential with spin-orbit coupling. Because

Of course, single-particle motion picture cannot describe the nuclear system perfectly. To describe the nuclear structure more precisely, the residual interaction, the deviation between single-particle interaction and many body correlation, needs to be taken account. In 1950s, nuclear physicists found one of the most important residual interaction, pairing correlation. One of the clear evidences for pairing correlation detected from experiment is the separation energy difference between odd nuclei and even nuclei. Fig. \ref{S_n} shows the separation energies for Sn, Sb, and Te isotopes as a function of neutron number. The isotopes in even neutron number have about $2\sim3$ MeV larger separation energies than the isotopes in odd neutron number.
It is odd-even staggering (OES) effect, which is detected in whole mass region of nuclei ( Fig. \ref{gap}). In this figure, $2\Delta$ corresponds to the separation energy difference between even-even nuclei and odd mass number nuclei. The approximate value is $2\Delta \approx 24A^{-1/2}$ MeV. It indicates a two-body attractive interaction should be included based on single-particle motion picture. In addition, if we examine the spin-parity of the ground states, we can find all of the even-even nuclei are $J^{\pi}=0^+$ states and most of the odd mass number nuclei have the same spin-parity as their single-particle orbital which the last nucleon fulfills in.
To explain these phenomena, it is natural to consider that the ground state of even-even nuclei consist of coupled $J^{\pi}=0^+$ pairs and the ground state of odd mass number nuclei consist of coupled $J^{\pi}=0^+$ pairs except one unpaired nucleon in last single-particle orbit. The two-body attractive interaction which couples two identical nucleons into $J^{\pi}=0^+$ state is called pairing correlation.
\begin{figure}[htbp]
 \begin{center}
    \includegraphics[width=100mm]{images/S_n.eps}
 \end{center}
  \caption{Separation energy for Sn, Sb, Te isotopes as a function of neutron number obtained from experimental data \cite{KJJ12}.}
  \label{S_n}
\end{figure}

\begin{figure}[htbp]
 \begin{center}
    \includegraphics[width=120mm]{images/gap.eps}
 \end{center}
  \caption{The odd-even mass differences $\Delta_n$, $\Delta_p$ for neutrons and protons (Zeldes {\it et al.} {\it Mat. Fys. Skr. Dan. Vid. Selsk.}, 3, no. 5 (1967)).}
  \label{gap}
\end{figure}

Because pairing correlation is two-body interaction, a theoretical framework beyond the mean field model seems to be necessary. However, there is a tricky approximated theory, BCS theory, enable us to describe pairing correlation within mean field model. In BCS theory, the ground state is superposed by the states with different particle number. It indicates that the ground state breaks symmetry in a space, namely U(1) gauge space. With symmetry breaking, the corresponding order parameter, pairing gap $\Delta$, has finite value. Based on BCS theory, because the energy of each particle in a $J^{\pi}=0^+$ pair is $E_k=\sqrt{(\epsilon_k-\lambda)^2+\Delta^2}$, where $\epsilon_k$ is single particle energy, $\lambda$ is Fermi energy, the minimum excitation energy is $2\Delta$. In fact, pairing gap corresponds to the separation energy difference between even-even nuclei and odd mass number nuclei discussed in previous paragraph. As same as superconductivity in condensed matter physics, the superfluid nucleus is defined when $\Delta\neq0$ in ground state. Thus, the OES detected from experiment caused from pairing correlation can be well explained by BCS theory.

After BCS theory was firstly introduced by J. Bardeen, L. Cooper, and J. R. Schrieffer in 1957, it was immediately generalized to Hartree-Fock-Bogoliubov (HFB) theory. HFB theory is the naturally extension of Hartree-Fock (HF) theory by introducing Bogoliubov transformation. Under Bogoliubov transformation, the particles become to quasiparticles, and the mean field potential is extended to generalized form including pairing field. After HFB theory had been applied to nuclear physics in 1958, the structure of ground states in nuclei was examined by plenty of studies. Up to now, HFB theory is the standard mean field theory to describe nuclear structure including pairing correlation.

% Recently, nuclear density functional theory has been developed based on HFB theory. and the computational codes are well developed. The properties of ground states in even-even nuclei are well studied by HFB theory.

%BCS theory is firstly introduced by J. Bardeen, L. Cooper, and J. R. Schrieffer in 1957 \cite{} to describe . BCS theory can describe the superconducting state or superfluid state caused by cooper pairs. In nuclear physics, the  created by pairing correlation correspond to the cooper pairs.

%  It indicates that the ground states consist of the $J^{\pi}=0^+$ pairs is more stable than other configurations. In addition, the ground states between even-even nuclei and neighborhood odd nuclei has large gaps of binding energy. In odd nuclei, the unpaired last neutron is the last single-particle level. The odd-even mass difference  To break the $J^{\pi}=0^+$ pair, we need large energy which

%Pairing correlation plays an important role near the Fermi energy.
%\begin{figure}[htbp]
% \begin{center}
%    \includegraphics[width=100mm]{images/Sn_isotope.eps}
% \end{center}
%  \caption{Low-lying excited states in Sn isotopes \cite{}. The absolute values of binding energy are adjusted at 0 MeV in the ground state of ${}^{116}$Sn.}
%  \label{Sn_isotope}
%\end{figure}

\subsection{Pairing collective motion}
\label{1-1-2}
Due to the superfluid state breaks the symmetry in U(1) gauge space, we expect that there are two types of elementary modes of excitation for pairing collective motion in nuclei. Two types of elementary modes of excitation are Nambu-Goldstone (NG) mode, pairing rotation, recovering the broken symmetry in U(1) gauge space, and Higgs mode, pairing vibration, fluctuating around the energy minimum point.

\begin{figure}[htbp]
 \begin{center}
    \includegraphics[width=100mm]{images/pair_coll.eps}
 \end{center}
  \caption{Schematic picture of pairing collective motion with superfluid ground state. Red  bi-direction arrow indicates pairing rotation, which rotates in $\Phi$ direction along the energy minimum circle degenerating at an absolute value of pairing gap $|\Delta|\neq 0$, in U(1) gauge space. Green bi-direction arrow indicates pairing vibration, which vibrates in $|\Delta|$ direction around one of the degenerated energy minimum points.}
  \label{Pair_coll}
\end{figure}
First, we discuss about pairing rotation. In classical picture, pairing rotation is the motion which the system rotates in the U(1) gauge space (red bi-direction arrow in Fig. \ref{pair_rot}). The motion is described by the coordinate, gauge angle $\Phi$, and the corresponding momentum, total particle number $N$. 
%It is similar to normal rotation in 3D space which described by the angles ($\theta_x, \theta_y, \theta_z$) and their conjugate angular momenta ($J_x, J_y, J_z$). 
Because the total particle number is unchanged in time evolution, the momentum $N$ is conserved quantity. In quantum picture, if we assume the ground state in a particle number $N_0$ as the ``ground state'', the ground states in $N=N_0\pm2, N_0\pm4,\cdots$ correspond to the ``excited states'' of pairing rotational mode. The excitation energy is $E_{\rm rot}^{\rm pair}=\frac{(N-N_0)^2}{2\mathcal{J}}$, where $\mathcal{J}$ is the pairing-rotational moment of inertia which characterizes the pairing rotation. In experiment, the pairing rotational band is expected to be observed in even-even isotope chain or even-even isotone chain between two magic numbers. One clear observation is found in Sn isotopes shown in Fig. \ref{pair_rot}. Except  a constant value and linear term with respect to mass number $A$, we can find the energies of ground states consist of a parabolic line - pairing rotational band. Recently, the systematic study for the pairing-rotational moment of inertia has been done in Ref. \cite{HN16}. The experimental moments of inertia obtained from two neutron(proton) separation energies have excellent correspondence with the theoretical moments of inertia obtained from NG mode in quasiparticle random phase approximation (QRPA) \cite{TV61} based on HFB mean field theory, not only for the spherical nuclei, but also for the deformed nuclei. Because the pairing-rotational moments of inertia are free from ambiguities caused from odd-even system, they are supposed to be excellent indicators of pairing correlation rather than pairing gaps.

Next, we discuss about pairing vibration. We emphasize that there are two types of pairing vibrations, one which vibrates around normal ground state ($\Delta\approx 0$), and another one which vibrates around superfluid ground state ($\Delta\neq 0$). The structures between two vibrational modes are quite different. 
%The former one is well understood because the structure of the excited $0^+$ states are $2np2nh$ configurations, detected from the energy spectra around ${}^{208}$Pb. 
Here, we only focus on the later one.
In classical picture, the later one vibrates around one of the degenerated energy minimum points. Depending on the intrinsic structure, various pairing vibrations can exist in one nucleus. 
%In general, the motion is supposed to be described by the coordinate, pairing gap $\Delta$, and its corresponding momentum\footnote{Because $\Delta$ is order parameter, it is naturally supposed to be the coordinate to describe the pairing vibration in previous studies. However, we found $\Delta$ is not proper coordinate in most of case from our study. The detail is discussed in Chapter 3 and 4.}.
In quantum picture, if we assume the ground state in a particle number $N_0$ as the ``ground state'', the excited states in $N=N_0, N_0\pm2, N_0\pm4,\cdots$ correspond to the ``excited states'' of pairing vibrational mode. The excitation energy is $E_{\rm vib}^{\rm pair}=n_{\alpha}\hbar\omega$, where $n_{\alpha}$ is the number of phonon and $\alpha$ is the label to distinguish different pairing vibrational modes. Because $2\Delta$ is the minimum energy to break one-pair, the collectivity is supposed to be strong when $E_{\rm vib}^{\rm pair}<2\Delta$. Thus, low-lying excited $0^+$ states in even-even nuclei are possible to be pairing vibrational states. Except the region around the magic number, excited $0^+$ states below $2\Delta$ exist in most of even-even nuclei. However, it is difficult to observe the signatures for pairing vibration directly from experiment. The properties of pairing vibration in nuclei are not well understood up to now.

The signatures for pairing collective motion can also be observed from the two-neutron(proton) transfer reaction \cite{BHR73}, such as $(p, t)$, $(t, p)$, and $({}^3{\rm He}, n)$. The cross sections of two particle transfer reaction strongly depend on the matrix elements of $J^{\pi}=0^+$ pair-transition operators (e.g. pair-additional transition: $\sigma( P_{ad};0_{\alpha}^+(N) \to 0_{\beta}^+(N+2))\sim |\braket{N+2,\beta|\sum_k a^{\dag}_k a^{\dag}_{\bar{k}}|N,\alpha}|^2$). From theoretical prediction, the pairing correlation is known to enhance $\sigma(g.s.\to g.s.)$ and decrease $\sigma(g.s.\to 0_{ex}^+)$ significantly, hence, $\sigma(g.s.\to 0_{ex}^+)/\sigma(g.s.\to g.s.)\ll 1$ is expected in superfluid system. One example of $(p, t)$ reactions is demonstrated in Fig. \ref{pair_rot}. As mentioned in previous paragraph, the ground states of even-even Sn isotopes consist of a pairing rotational band. In such situation, we can find intraband transition is dominant with respect to interband transition. Therefore, the cross sections of two-neutron transfer reaction is consistent with theoretical prediction. In general, the ratio $\sigma(g.s.\to 0_{ex}^+)/\sigma(g.s.\to g.s.)$ for two-neutron(proton) transfer reaction are excellent indicators to describe the collectivity of pairing dynamics.

\begin{figure}[htbp]
 \begin{center}
    \includegraphics[width=100mm]{images/pair_rot.eps}
 \end{center}
  \caption{Experimental energies of the $J^{\pi}=0^+$ states of the Sn isotopes (ground state and pairing vibration),
populated in $(p, t)$ reactions. The heavy drawn horizontal lines
represent the values of the expression $E=-B({}_{50}^{A}{\rm Sn}_{N})+E_{exc}+
8.124A+46.33$ MeV, where $B({}_{50}^{A}{\rm Sn}_{N})$ is the binding energy
of the Sn isotopes of mass number $A$, and $E_{exc}$ is the
weighted [with $\sigma(0_i^+)$] average energy of the excited $0^+$ states
below 3 MeV. The dashed and dotted lines represent the parabolas
given in the insets, corresponding to the ground state and to
the (average) excited state-based pairing rotational bands. The
absolute values of the $g.s.\to g.s.$ integrated cross sections (in $\mu$b units) are given (perpendicular) to the abscissa, as a function
of $N$. The values in the shaded areas correspond to the experimental values, while the remaining values correspond to theoretical
predictions. For the first group (experimental), the relative
(in \%)$(p, t)$, pairing vibrational, cross sections $\sum_i\sigma(g.s.\to 0_i^+)$
normalized with respect to the ground state cross sections are in parenthesis. The detailed information are written in \cite{PF11}.}
  \label{pair_rot}
\end{figure}

\section{Motivation of this study}

\subsection{Mysterious excited $0^+$ states and pairing dynamics}
In most of nuclei, the structure of excited $0^+$ states is complicate and unclear. The reason is that not only pairing correlation but also deformations and multipole correlations influence the excited $0^+$ states significantly. The various correlations competing with each other construct plenty of mysterious excited $0^+$ states.

After the concept of pairing vibration was introduced, it encountered the problem for the large asymmetry cross sections between $(p, t)$ and $(t, p)$ reaction in actinide region observed from experiment. In low-lying energy region, the cross section of $\sigma(g.s.\to 0_{ex}^+)$ is about $15\%$ of $\sigma(g.s.\to g.s.)$ for $(p, t)$ reaction \cite{MEFSS70, MEFSS72}, while the signal of $\sigma(g.s.\to 0_{ex}^+)$ is extremely weak for $(t, p)$ reaction \cite{Cas72}. To explain such phenomena, the coexistence of both oblate shape and prolate shape in one state, and the mixing of pairing correlations between oblate and prolate shape was suggested \cite{GJV71,RK72}. Another interpretation for such phenomena is using microscopic interaction, including multipole pairing correlations and quadrupole particle-hole interactions \cite{RB76}. Although the approaches are different, they succeeded to reproduce the asymmetry cross section of two-neutron transfer reaction in actinide region.

$\beta$ vibration
Collective motion is the essential idea to explain the excited states in nuclei. The quadrupole collective motion described by the quadrupole deformation parameters $\alpha_{2\mu} (\mu=\pm2,\pm1,0)$ is known as the most typical collective motion. In the framework of quadrupole collective motion, the low-lying excited $0^+$ states can be roughly categorized into two types: quadrupole vibration emerging above two-phonon excitation in spherical nuclei, and $\beta$ vibration emerging in quadrupole deformed nuclei. The excitation energies and $B(E2)$ transition strength is well understood under harmonic approximation.

\subsection{Theoretical problem}
%As discussed in previous section, the low-lying excited $0^+$ states are not be understood in general. 
We aim to understand the mysterious excited $0^+$ states discussed in previous section based on mean field approach. In low-energy region, the pairing correlation supposed to be important as the quadrupole correlation. The most well-known and successful model is Bohr model, which was introduced to describe low-energy nuclear collective motion in quadrupole degrees of freedom with the deformation parameters $(\beta,\gamma)$ and the Euler angles $(\phi,\theta,\psi)$. However, the pairing degrees of freedom is not included in the Bohr model. As discussed in Sec. \ref{1-1-2}, the pairing vibrational degrees of freedom supposed to be a key ingredient to understand the low-lying  excited $0^+$ states.
 the pairing vibrational degrees of freedom is necessary to include in the collective model. the role of pairing correlation is hardly discussed up to now.
The degrees of freedom for pairing collective motion are supposed to be necessary including the collective model.

The study about the collective model for pairing collective motion was firstly introduced by Bes. et al. \cite{BBPK70}. They constructed phenomenological pairing collective model  by two degrees of freedom, the pair deformation parameter (pairing gap) $\Delta$ and the global gauge angle $\Phi$. After the pairing collective model was formulated, it was applied to several studies for nuclear structure \cite{GPBW85, ZPPRS99, P07}. 
In Ref. \cite{ZPPRS99}, the $B(E2)$ transition improved by introducing pairing degrees of freedom. On the other hand, the wave function of ground state from collective model looks like unusual because the pairing gap $\Delta_{vib}$ at the peak of the wave function is smaller than the pairing gap $\Delta_{eq}$ obtained from BCS equation.

It is supposed to be important to find the collective coordinates describing the pairing collective motions from the microscopic degrees of freedom without any assumption.

The difficult point is the requantization of the pairing collective motion. Such difficulty is not found in quadrupole collective motion.

\subsection{Our idea}

\subsection{Outline of this thesis}
In this study, we concentrate in the pairing model which only contains pairing interaction in two-body interaction to construct the theoretical framework of the requantization.

@@

To examine the pairing dynamics influencing the structure of nuclei, we can only concentrate the nucleons near the Fermi energy. The Hamiltonian of pairing model is
\begin{align}
	H &= \sum_{l=1}^L \epsilon_l n_l - g \sum_{l,l'} S_l^+ S_{l'}^- ,
\end{align}
where
\begin{align}
	n_l &= \sum_m a^{\dag}_{lm}a_{lm} \\
        S_l^{+} &= \sum_{m>0}a_{lm}^{\dag}a_{l\overline{m}}^{\dag} ,
\quad   S_l^{-} = S_l^{+\dag} .
\label{S_+}
\end{align}
In $L$ single-particle level system, each single-particle energy $\epsilon_l$ possesses $(2\Omega_l)$-fold
degeneracy ($\Omega_l=j_l+1/2$)
and $\sum_{m>0}$ indicates the summation over $m=1/2,3/2,\cdots,$ and $\Omega_l-1/2$. $\overline{m}$ indicates the time reversal quantum number with respect to $m$. The strength of pairing correlation is represented by coupling constant $g$. We also define a seniority $\nu_l$, which is the number of unpaired particles in each level. The total number of unpaired particles is $\nu=\sum_l\nu_l$. The unpaired particles have no pairing interaction, and play a part in Pauli blocking effect.\par
In one-level system ($L=1$), the pairing dynamics is well understood. The analytic solution of the total energy is
\begin{align}
	E_{\nu}(N) = \epsilon N - \frac{1}{4}g(N-\nu)(2\Omega-N-\nu+2) .
	\label{onelevel}
\end{align}
The pairing rotational band created by the pairing correlation is completely parabolic with the moment of inertia $2/g$. Also, the excited states in a specified $N$ is only described by the quantum number $\nu$. For example, the first excited state in even particle system is $\nu=2$ state, which correspond to one $J^{\pi}=0^+$ pair broken state.\par
The profound dynamics occurs in multi-level pairing model ($L\ge 2$). The pairing rotational mode becomes complex due to the shell effect. In addition, several pairing vibrational modes emerge. The pairing vibration is not single particle motion specified by $\nu_l$, but a collective motion with the excitation of $J^{\pi}=0^+$ pairs. In this chapter, we explain the numerical exact solution and derive the TDHFB dynamics in multi-level pairing Hamiltonian.

@@

\subsection{Richardson equation}
As mentioned above, diagonalizing the Hamiltonian is simple method to obtain the exact solution.
However, diagonalization is time consuming when the model space becomes large. The calculation time is proportional to the cubes of the dimension of the model space For example, there are 16 single particle levels between the magic number 50 and 82 in deformed nuclei. Because the dimension for the system is $D=2^{16}=65536$, the calculation time in deformed system is $\sim 10^5$ times as much as the case in spherical system. \par
The alternative method to obtain the exact solution is to solve Richardson equation. The idea was introduced by R. W. Richardson \cite{Richardson1, Richardson2, Richardson3}. We divide the total energy into two parts
\begin{align}
  E &= \braket{\psi|H|\psi} \\
  &= \sum_l \epsilon_l \nu_l + \sum_{\alpha=1}^M E_{\alpha} .
\end{align}
The first term is the energy of unpaired particle, and the second term is the total energy of $J^{\pi}=0^+$ pairs. The pair-energy $E_{\alpha}$ is derived from the non-linear equations
\begin{equation}
  1 - g\sum_l \frac{\Omega_l-\nu_l}{2\epsilon_l-E_{\alpha}} + 2g\sum_{\beta(\neq\alpha)}^M \frac{1}{E_{\beta}-E_{\alpha}} = 0 .
\end{equation}
$M$ indicates the number of $J^{\pi}=0^+$ pairs in the system, and equals to the numbers of the non-linear equations. $E_{\alpha}$ can become not only real numbers, but also complex conjugate pairs when $M\ge 2$.

\section{TDHFB dynamics}
%In realistic nuclear system with nuclear effective interaction, the necessary model space is huge. To diagonalize the effective Hamiltonian is extremely difficult task. Such study is done by shell model calculation. In alternative way, we only prepare a limited model space, which supposed to be reproduce the important parts of nuclear properties. The time-dependent mean-field theory 
To understand the classical picture of the pairing dynamics, time-dependent mean-field (TDMF) theory is useful tool. With pairing correlation, TDMF theory attributes to TDHFB theory.
The TDHFB equation can be derived from the time-dependent variational principle,
\begin{equation}
  \delta\mathcal{S}=0, \quad 
  \mathcal{S} \equiv \int \braket{\phi(t)|i\frac{\partial}{\partial t}-H|\phi(t)}dt ,
  \label{variation}
\end{equation}
where $\ket{\phi(t)}$ is the time-dependent generalized Slater determinant (coherent state) obtained from Thouless' theorem
\begin{equation}
  \ket{\phi(t)} = \mathcal{R}\exp{\left( \sum_{k<k'}Z_{kk'}(t)\beta_k^{\dag}\beta_{k'}^{\dag}
    \right)}\ket{\Phi_0} .
\end{equation}
$\beta_k^{\dag}$ is quasiparticle creation operator and $\mathcal{R}$ is normalization factor. $\ket{\Phi_0}$ is generally chosen as ground state or vacuum state. $Z_{kk'}(t)$ is time-dependent complex number determined by TDHFB equation.

@@

\subsection{Constant of motion}
\label{constant}
The number of the TDHFB degrees of freedom is same as the number of the single-particle levels. However, because the Hamiltonian depends only on the relative difference in the gauge angles, the global gauge angle 
\begin{align}
	\Phi = \frac{1}{L}\sum_l \chi_l
\end{align}
is a cyclic variable. With the definition for relative gauge angle (e.g. $\phi_l = \chi_l-\chi_1$), we can find that the conjugate momentum of $\Phi$ is the sum of $j_l$
\begin{align}
	J \equiv \frac{\partial\mathcal{L}}{\partial\dot{\Phi}} = \sum_l j_l .
\end{align}
From (\ref{occ}), $J$ is related to the total particle number
\begin{align}
	N = \nu + 2J .
\end{align}
The Hamilton's equation for $J$
\begin{align}
	\dot{J} = -\frac{\partial\mathcal{H}}{\partial \dot{\Phi}} = 0, \quad J = {\rm const}
\end{align}
indicates that the total particle number is conserved in time evolution.
The canonical variable $(\Phi,J)$ describes the pairing rotation. The pairing rotational band emerges in the particle number chain ($\cdots,N-2,N,N+2,\cdots$). The last term in (\ref{Hamiltonian}) contains the higher order terms with respect to $N$. It related to the shell effect. In single-level system, the last term in (\ref{Hamiltonian}) disappears and (\ref{Hamiltonian}) attributes to (\ref{onelevel}). The pairing rotational band is exactly parabolic. In multi-level system, the coupling constant $g$ and single-particle energy $\epsilon_l$ affects significantly to the last term in (\ref{Hamiltonian}). The parabolic structure can be broken when the shell effect is strong.

@@

\section{Theoretical framework of ASCC+SPA in non-integrable system}

@@

\section{Matrix elements of the spin-coherent state}
\label{formula1}
The time-dependent coherent state classified into the spin-coherent state is employed in (\ref{coherent}). We derive the necessary formula for the spin-coherent state in this subsection. \par
First, we consider the single-level case.
With the commutation relation in (\ref{SU2commu}), the SU(2) quasispin operators fulfill the following relations
\begin{align}
  \hat{S}^{\pm}\ket{S,S^0} &= \sqrt{(S\mp S^0)(S\pm S^0+1)} \ket{S,S^0\pm 1}
  \label{S_pm} \\
  \hat{S}^0\ket{S,S^0} &= S^0\ket{S,S^0} .
  \label{S_0}
\end{align}
Using above relations, the coherent state can be expanded as 
\begin{align}
  \ket{Z} &= (1+|Z|^2)^{-S}e^{Z \hat{S}^+}\ket{S,S^0=-S} \nonumber \\
  &= (1+|Z|^2)^{-S}\sum_{n=0}^{2S}\sqrt{\frac{(2S)!}{n!(2S-n)!}}Z^n\ket{S,-S+n} .
\end{align}
We can find the normalization condition is fulfilled
\begin{align}
  \braket{Z|Z} &= (1+|Z|^2)^{-2S}\sum_{n=0}^{2S}\frac{(2S)!}{n!(2S-n)!}|Z|^{2n} \nonumber \\
  &= (1+|Z|^2)^{-2S}\sum_{n=0}^{2S} {}_{2S}\mathrm{C}_n(|Z|^2)^n \nonumber \\
  &= (1+|Z|^2)^{-2S}\times(1+|Z|^2)^{2S} \nonumber \\
  &= 1 .
\end{align}
The overlap is
\begin{align}
  \braket{\eta|Z} &= \frac{(1+\eta^*Z)^{2S}}{(1+|\eta|^2)^S(1+|Z|^2)^S} .
\end{align}
